---
title: "autocorrelation"
author: "Aske Svane Qvist"
date: "25/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Autocorrelation

## Question
Is the population of single women in Denmark spatially correlated? What is the correlation and how significant is the trend?


### Import packages
```{r}
library(raster)
library(rgeos)
library(sf)
library(tidyverse)
library(htmltools)
library(googlesheets4)
library(mapview)

```


### Get spatial map of Denmark with municipalities
```{r}
# Load the spatial data
mun_sp <- getData('GADM', country = 'DK', level = 2)

# change to sf object
mun_sf <- st_as_sf(mun_sp)

# change UTM to 32632
mun <- st_transform(mun_sf, crs = 32632)
mapview(mun)

# Straighten the names
sort(mun$NAME_2)
which(grepl("Å",mun$NAME_2))
which(grepl("Vest",mun$NAME_2))

mun$NAME_2[31] <- "Aarhus"
mun$NAME_2[21] <- "Høje-Taastrup"
mun$NAME_2[60] <- "Vesthimmerlands"

```


### Import data of single men/women in Danish municipalities and tidy up the data

Import and select relevant columns for further analysis
```{r}
# Import csv
data <- read_csv("data/DKStatData_CivilStatus2015_2020.csv")

# Quick overview
head(data)
summary(data)

# Clean up the data
data <- data %>% 
  
  # Change character columns to factor
  mutate(
    Status = as.factor(Status),
    Sex = as.factor(Sex),
    Region = as.factor(Region)) %>% 
  
  # Rename year columns to actual years
  rename(
    year2020 = Y2020K1) %>% 
  
  filter(
    Sex == "Kvinder") %>% 
  
  drop_na()

# Select relevant columns
data <- data[1:4]

```

More data tidying
```{r}
# check levels in column status
data %>% group_by(Status) %>% summarise()

# Create single dummy variable (1 = single, 0 = not single)
data$single <- ifelse(data$Status == "Gift/separeret", "0", "1")

# Find the sum of ALL women in each municipality
data_sum <- data %>% group_by(Region) %>% summarise(
  total = sum(year2020)
)

# Merge the data with the sum data
data <- merge(data, data_sum, by="Region")

# I find the amount of single women in each municipality by calculating the difference between married (dummy 0) and total amount of women.
# I use the number of singles and total to calculate a ratio
data <- data %>% 
  
  filter(single == "0") %>% # We only need the rows with the married women including the total for each mun.
  
  mutate(
    singles = total - year2020, # number of singles in the municipality
    single_ratio = round((total - year2020) / total,2)) %>% # the ratio of single women
  
  subset(select = -c(Status, year2020, single)) # drop columns

# Merge with geometry dataframe
dataGeo <- mun %>% subset(select = "NAME_2") %>% merge(data, by.x = "NAME_2", by.y = "Region") %>% rename(Municipality = NAME_2)

```

Quick plot
```{r}
# Using mapview to get a quick look at the difference between municipalities
dataGeo %>% 
  group_by(Municipality) %>% 
  dplyr::select(single_ratio) %>% 
  mapview()

```


### Run autocorrelation analysis

Make the data ready
```{r}
# Quick overview
plot(dataGeo$geometry)

# I simplify the geometry a bit
mun_sm <- st_cast(st_simplify(dataGeo, dTolerance = 250),
                     to = "MULTIPOLYGON")

plot(mun_sm$geometry) # not really visible


# Use the spdep package
library(spdep)

# Make neighbor list following queen adjacency
nb <- poly2nb(mun_sm$geometry)
nb

```

Plot what we define as neighbors according to queen adjacency
(Neighbors are all given municipalities sharing borders or a corners)
```{r}
# Get center points of each municipality
mun_centers <- st_coordinates(st_centroid(mun_sm$geometry))

# Show the connections
plot(mun_sm$geometry); plot(nb, mun_centers, col = "red",add = TRUE)
```

Run Moran's I autocorrelation test
```{r}
# Run a Moran I test on the proportion of single women
moran.test(mun_sm$single_ratio, 
           nb2listw(nb, style = "W",zero.policy=TRUE),
           zero.policy=TRUE)


# Do a Monte Carlo simulation to get a better p-value
moran.mc(mun_sm$single_ratio,nb2listw(nb, zero.policy=TRUE),zero.policy=TRUE, nsim = 999)

```

Both the regular statistical test as well as the Monte Carlo simulation show a correlation of 0.28. The p-value is 0.005 thus highly significant. In sum, the ratio of single women in each municipality is geographically correlated.

